---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## 
```{r}
url <- "https://tinyurl.com/COVID-2019"

virus <- read.csv(url)

tail(virus)
```


## Q1. How many total infected cases are there around the world?
#check in cases and sum it up

```{r}
sum(virus$cases)
```


## Q2. How many deaths linked to infected cases have there been?


```{r}
virus$type
```

```{r}
table(virus$type)
```

```{r}
inds <- virus$type == "death"

##use square brackets to print out all the ones that have high death

virus[inds,]
```

```{r}
sum(virus[inds, "cases"])

total_cases <- sum(virus$cases)
```



## Q3. What is the overall death rate?

```{r}
death_cases <- sum(virus[inds, "cases"])
death_cases
```

```{r}
round(death_cases/total_cases *100, 2)
```


## Q4. What is the death rate in China?

```{r}
table(virus$Country.Region)
```

```{r}
virus$Country.Region == "Mainland China"
```

```{r}
sum(virus$Country.Region == "Mainland China")
```

## Make `china_total` a vector
```{r}
China_total <- sum(virus$Country.Region == "Mainland China")
```

```{r}
China_total
```


## make `China_deaths` vectors 

```{r}

china_deaths <- sum(virus$Country.Region == "Mainland China"&virus$type == "death")

china_deaths
```


```{r}
round(china_deaths/China_total * 100, 2)
```





## Q5. What is the death rate in Italy, Iran and the US?

##italy 
```{r}
sum(virus$Country.Region == "Italy")
```

```{r}
Italy_total <- sum(virus$Country.Region == "Italy")

Italy_total
```
```{r}
Italy_deaths <- sum(virus$Country.Region == "Italy"&virus$type == "death")

Italy_deaths
```

```{r}
round(Italy_deaths/Italy_total * 100, 2)
```


##Iran

```{r}
sum(virus$Country.Region == "Iran")
```

```{r}
Iran_total <- sum(virus$Country.Region == "Iran")

Iran_total
```

```{r}
Iran_deaths <- sum(virus$Country.Region == "Iran"&virus$type == "death")

Iran_deaths
```


```{r}
round(Iran_deaths/Iran_total * 100, 2)
```


#US

```{r}
sum(virus$Country.Region == "US")
```




```{r}
round(72/total_cases * 100, 2)
```

```{r}
US_total <- sum(virus$Country.Region == "US")

US_total
```

```{r}
US_deaths <- sum(virus$Country.Region == "US"&virus$type == "death")

US_deaths
```

```{r}
round(US_deaths/US_total * 100, 2)
```

