---
title: "Corona virus hw"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## 
```{r}
url <- "https://tinyurl.com/COVID-2019"

virus <- read.csv(url)

tail(virus)
```
```{r}
table(virus$cases)
```


## Q1. How many total infected cases are there around the world?
#check in cases and sum it up

```{r}
sum(virus$cases)
```


## Q2. How many deaths linked to infected cases have there been?


```{r}
virus$type
```

```{r}
table(virus$type)
```

```{r}
inds <- virus$type == "death"

##use square brackets to print out all the ones that have high death

virus[inds,]
```

```{r}
sum(virus[inds, "cases"])
```

```{r}
total_cases <- sum(virus$cases)
```


## Q3. What is the overall death rate?


inds <- virus$type == "death"

```{r}
death_cases <- sum(virus[inds, "cases"])
death_cases
```



```{r}
round(death_cases/total_cases *100, 2)
```


## Q4. What is the death rate in China?



```{r}
table(virus$Country.Region)
```

```{r}
virus$Country.Region == "Mainland China"
```

## number of total cases in `Mainland.China`
```{r}
sum(virus[virus$Country.Region == "Mainland China", "cases"])

China_total <- sum(virus[virus$Country.Region == "Mainland China", "cases"])
```

## number of total deaths in `Mainland.China`

```{r}
library(dplyr)

dplyr::filter(virus, Country.Region == "Mainland China", virus$type == "death")
```

```{r}
virus[virus$Country.Region == "Mainland China" & virus$type == "death", "cases"]
```

```{r}
sum(virus[virus$Country.Region == "Mainland China" & virus$type == "death", "cases"])

China_deaths <- sum(virus[virus$Country.Region == "Mainland China" & virus$type == "death", "cases"])
```


## rate of `Mainland.China` deaths

```{r}
round(China_deaths/China_total * 100, 2)
```


## Q5. What is the death rate in Italy, Iran and the US?

##italy

## number of deaths in Italy
```{r}
sum(virus[virus$Country.Region == "Italy" & virus$type == "death", "cases"])

Italy_deaths <- sum(virus[virus$Country.Region == "Italy" & virus$type == "death", "cases"])
```



##use this for total cases
```{r}
virus[virus$Country.Region == "Italy", "cases"]
```
##use this for total cases
```{r}
sum(virus[virus$Country.Region == "Italy", "cases"])

Italy_total <- sum(virus[virus$Country.Region == "Italy", "cases"])
```


```{r}
round(Italy_deaths/Italy_total * 100, 2)
```


##Iran

## number of deaths in Iran 

```{r}
sum(virus[virus$Country.Region == "Iran" & virus$type == "death", "cases"])

Iran_deaths <- sum(virus[virus$Country.Region == "Iran" & virus$type == "death", "cases"])
```




##use this for total cases
```{r}
sum(virus[virus$Country.Region == "Iran", "cases"])

Iran_total <- sum(virus[virus$Country.Region == "Iran", "cases"])
```



```{r}
round(Iran_deaths/Iran_total * 100, 2)
```



#US

```{r}
sum(virus[virus$Country.Region == "US" & virus$type == "death", "cases"])

US_deaths <- sum(virus[virus$Country.Region == "US" & virus$type == "death", "cases"])
```


##use this for total cases
```{r}
sum(virus[virus$Country.Region == "US", "cases"])

US_total <- sum(virus[virus$Country.Region == "US", "cases"])
```


```{r}
round(US_deaths/US_total * 100, 2)
```

