---
title: "Structure based drug design"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

#Download "1hsg" form PDB into the RStudio Project directoty, then cleanup the HIV-Pr. We will make a separate set of "protein-only" and "ligand-only" PDB files

```{r}
library(bio3d)
```

```{r}
file.name <- get.pdb("1hsg")
```

##now the file form PDB ("1hsg") is stored as a file to the right

#We will use `read.pdb()`, `atom.select()`, and `write.pdb()` functions to make our seperate "protein-only" and "ligand-only" PDB files

```{r}
read.pdb(file.name)
```

```{r}
hiv <- read.pdb(file.name)
hiv
```

```{r}
prot <- atom.select(hiv, "protein", value=TRUE)
```
```{r}
lig <- atom.select(hiv, "ligand", value=TRUE)
```

##review whether you have MK1 as you ligand

```{r}
lig
```

##now review the prot

```{r}
prot
```


#This will give you files on the right hand side. One file each for protein and for ligand

```{r}
write.pdb(prot, file="1hsg_protein.pdb")
```


```{r}
write.pdb(lig, file="1hsg_ligand.pdb")
```

##you can do this process on anything 

#now switch over to AutoDock Tools(ADT) from teh MGLTools you just downloaded


##make sure to get the files that have the protein charges into this folder

#go to the terminal and type in ` "\Program Files (x86)\The Scripps Research Institute\Vina\vina.exe" --help` in terminal 

#no type this in 
```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")

```

##Read the original ligand with added hydrogens that you produced earlier and use the `rmsd()` function to compare to your docking results

```{r}
res2 <- read.pdb("all.pdbqt", multi=TRUE)
```
```{r}
ori <- read.pdb("ligand.pdbqt")
```

```{r}
rmsd(ori, res2)
```

```{r}
res2
```

```{r}
pdb <- read.pdb("1hel")
modes <- nma( pdb )
m7 <- mktrj(modes, mode=7, file="mode_7.pdb")
```

```{r}
library(DESeq2)
```


