---
title: "Class14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

head(mxl)
```

##we want to look at the 2nd column that contains the genotype information

```{r}
mxl$Genotype..forward.strand.
```

```{r}
table(mxl$Genotype..forward.strand.)
```

##RNA-seq result analysis for different genotypes of this SNP



```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")

head(expr)
```

```{r}
expr$geno == "G/G"
```

```{r}
x <- 1:10
x[ x > 5 ]
```

```{r}
expr$geno[expr$geno == "G/G"]
```

```{r}
expr[expr$geno == "G/G", ]
```

```{r}
summary(expr[expr$geno == "G/G", ]$exp)
```

```{r}
summary(expr[expr$geno == "A/G", ]$exp)
```


```{r}
summary(expr[expr$geno == "A/A", ]$exp)
```


```{r}
hist(expr[expr$geno == "G/G", ]$exp, breaks =20)
```

##use the boxplot function. the input data will be `expr()`. how do we draw a useful plot ... boxplot

```{r}
?boxplot()
```

```{r}
boxplot( exp ~ geno, data=expr)
```

##add notch

```{r}
boxplot( exp ~ geno, data=expr, notch=TRUE)
```



#how many samples are we looking for?

```{r}
nrow(expr)
```


