---
title: "Class 9 updated"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##K-means clustering

Let's try the kmeans() function in R to cluster some made-up example data

#tmp
##c puts things into a vector
##first call to R gives me 30 points with a mean of -3 (30 points clustered around -3)
##second call to R gives me 30 points with a mean of 3 (30 points clustered around 3)

##cbind binds data?? and will put x as first column, and y as a second column 
##rev means list the numbers in reverse

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))

x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```

#KM is KMEANSs(x, CENTERS=2, NSTART=20)
```{r}
km <- kmeans (x, centers =2, nstart=20)
```
```{r}
print(km)
```

##clustering vectors - the different clusters (cluster 1 and cluster 2)

##What is the output object 'km' ? 
##Atributes function to find this information use 'attributes()'

```{r}
attributes(km)
```

##size gives you the number of points for vector 1 and 2
```{r}
km$size
```


#cluster is a vector of integers indicating the cluster to which each point is allocated

```{r}
km$cluster
```

##let's check how many 2s and 1s are in this vector with the cluster function
```{r}
table(km$cluster)
```


##plot x colored by the kmenas cluster assignment and add cluster centers as blue points
```{r}
plot(x, col=km$cluster)
```

```{r}
c(rep("red", 30), rep("blue", 30))
```

```{r}
plot(x, col=km$cluster+1)
```


```{r}
plot(x, col=km$cluster+2)
```


```{r}
plot(x, col=km$cluster)
points(km$centers, col="blue")
```

##pch is for characters so you can get a big blue square

```{r}
plot(x, col=km$cluster)
points(km$centers, col="blue", pch=15)
```

#can change shape of pch if you assign it a different number

```{r}
plot(x, col=km$cluster)
points(km$centers, col="blue", pch=17)
```

#dist to calculate the distance
##HCLUST(DIST(X)) calcualting distance matrix

#the hclust() function is the main hierarchcal clustering method in R and it **must** be passed a **distance matrix** as input not your raw data!

#you can call seqeunce identities here, or structures
```{r}
hc <- hclust(dist(x))
```

```{r}
hc
```

```{r}
plot(hc)
```

##1st vector contains points 1-30, 2nd vector contains points 31 - 60

#cut below red line and get two branches

```{r}
plot(hc)
abline(h=6, col="red")
```

#tell it a height of where to cut (cut by height **h**)
#lty to get a dashed line

```{r}
plot(hc)
abline(h=6, col="red", lty=2)
cutree(hc, h=6)
```

##Now cut them into kgroups
#the higher the k gropu number, the higher the number of groups you get
```{r}
plot(hc)
abline(h=6, col="red", lty=2)
abline(h=6, col="blue", lty=2)
cutree(hc, k=3.5) # Cut into k grps
```

##use a table to see how many points in each group
```{r}
table(cutree(hc, h=3.5))
```


##ask for k instead; you can also ask `cutree()` for the `k` number of gorups that you want

```{r}
cutree(hc, k=5)
```

```{r}
cutree(hc, k=3.5)
```

##Using different hierarchical clustering methods

hc.complete <- hclust(d, method="complete")
hc.average <- hclust(d, method="average")
hc.single <- hclust(d, method="single")

# Step 1. Generate some example data for clustering
```{r}
x <- rbind(
matrix(rnorm(100, mean=0, sd=0.3), ncol = 2), # c1
matrix(rnorm(100, mean=1, sd=0.3), ncol = 2), # c2
matrix(c(rnorm(50, mean=1, sd=0.3), # c3
rnorm(50, mean=0, sd=0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```


# Step 2. Plot the data without clustering
```{r}
plot(x)
```

# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
```{r}
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```


#Q. Use the dist(), hclust(), plot() and cutree()
functions to return 2 and 3 clusters

#membership factor (grps3)

#Q. How does this compare to your known 'col' groups?

```{r}
hc <- hclust(dist(x))
plot(hc)
```
```{r}
grps3 <- cutree(hc, k=3)
grps3
```

```{r}
table(grps3)
```

#color it by the membership (grps3) factor
```{r}
plot(x, col=grps3)
```

```{r}
table(grps3)
```

##set up a cross table

##this will show you the points that were false-positives (the smallest number in the table per c1, c2, or c3)
```{r}
table(grps3, col)
```
#**this should have given me a tale c1, c2, c3 rows and 1, 2, 3 columns**

#Principal Component Analysis (PCA) the absolute basics
#when you have a tone of cell lines
#*PCA converst correlation or lack there of a represenation that you can actually look at* PCA shows us that those that cluster together are similar

#*the x-axis (PC1) are more important*


#the main function in bas R for PCA is called prcomp(). Here we will use PCA to examine the funny food that folks eat in the UK and N. Ireland


```{r}
x <- read.csv("UK_foods.csv")
```



## Preview the first 6 rows
```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```

spotting major differences in trends
```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```



```{r}
pairs(x, col=rainbow(10), pch=16)
```

# Use the prcomp() PCA function 
##t means transpose (from math)
```{r}
pca <- prcomp(t(x))
```

#gives the standard deviation
#looking at portion of variance. look at most amount of variance
```{r}
summary(pca)
```

##second and third row from above are in percentages

#look at the arributes



# Plot PC1 vs PC2
```{r}
plot(pca$x[,], pca$x[,], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x)) #this adds text for x and y coordinates and the country names
```

#Customize your plot so that the colors of the country names match the colors in our UK and Ireland map and table at start of this document

```{r}
plot(pca$x[,], pca$x[,], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), col=c("black", "red", "blue", "dark green")) 

#adding colors: black for england b/c first on the table, wales in red, scotland in blue, and ireland in dark green
```

#gives me a barplot
```{r}
plot(pca)
```

## Lets focus on PC1 as it accounts for > 90% of variance 
```{r}

par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,2], las=2 )
```


```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,3], las=2 )
```


```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=1 )
```

#las gives you just one word and changes the y-axis

```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=3 )
```







