---
title: "class10"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##PRCOMP(x, SCALE = T)

##get our input date. our data for today comes from Winsconsin Breat 

```{r}
wisc.df <- read.csv("WisconsinCancer.csv")
head(wisc.df)
```

#clean data. select from column 3 to 32

```{r}
wisc.data <- as.matrix(wisc.df[ , 3:32])
head(wisc.data)
```

#Q. How many patients are there in this dataset?

```{r}
nrow(wisc.df)
```


#Q. How many cancere and non-cancer patients?
access with dollar sign

```{r}
wisc.df$diagnosis
```

#use the table version

```{r}
table(wisc.df$diagnosis)
```
#test

```{r}
table(wisc.df$diagnosis == "M")
```


#use a graph to see how many of the columns have "mean" component 
##graphs look for patterns within a vector

##**use colnames()

```{r}
colnames(wisc.df)
```

##use *grep()*, search in the help box,  and use the ones without the equal sign
```{r}
grep("_mean", colnames(wisc.df))
```

##to just get the columns with the **_mean**
```{r}
grep("_mean", colnames(wisc.df), value = TRUE)
```

#the number of columns there are with *_mean*
```{r}
length(grep("_mean", colnames(wisc.df)))
```

##Enter Principal of Component Analysis

##First check whether our input should be scaled. let's check the sd() and mean() of all our columns in `wisc.data`

##APPLY(x, 2,  function you want to apply like Standard deviation)

```{r}
apply(wisc.data, 2, sd)
```

#try to round this up. use "2" to command it to show 2 decimal points

```{r}
round(apply(wisc.data, 2, sd), 2)
```

#perform PCA on wisc.data by completing the following

```{r}
wisc.pr <- prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```

```{r}
plot(wisc.data)
```

#this is a hot mess! we need to cook our own PCA plot. need to access the results within the `wisc.pr` 

```{r}
attributes(wisc.pr)
```

##want ot use `$x` component to make our PCA. plot column PC1 to PC2

```{r}
plot(wisc.pr$x[ ,1:2])
```



```{r}
plot(wisc.pr$x[ ,1], wisc.pr$x[ , 2], col=wisc.df$diagnosis)
```

```{r}
table(wisc.df$diagnosis)
```

#this is super important to do clusters
```{r}
hc <- hclust (dist(wisc.data))
plot(hc)
```

#use result of PCA fo do my clustering

```{r}
hclust(dist(wisc.pr$x[ , 1:3]))
```



```{r}
wisc.pr.hc <- hclust(dist(wisc.pr$x[ , 1:3]), method="ward.D2")
plot(wisc.pr.hc)
```

#where to cut this clustogram. need the `cutree()`

```{r}
grps <- cutree(wisc.pr.hc, k=2)
grps
```

#how many 1s and 2s? use a table

```{r}
table(grps)
```

#ho wmnay are benign (B) and how many are malignant (M)


```{r}
table(grps, wisc.df$diagnosis)
```

#24 and 33 are false positives

#Now look at NEW DATA FILE

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```





```{r}
plot(wisc.pr$x[ ,1], wisc.pr$x[ , 2], col=wisc.df$diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```


